---
title: "Visualising and Testing Data"
author: "Sam Forbes"
date: "4 April 2018"
output: 
  ioslides_presentation:
    highlight: 'espresso'
    fig_height: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, eval = T, warning = F, prompt = T, comment = '')
```

## Load data
```{r}
setwd('//ueahome/eresssf5/gax18aku/data/Documents/R_Workshops')
choc <- read.csv('choccies.csv', header = T, sep = ',')
library(dplyr)
library(reshape2)
```
```{r, echo =F}
library(knitr)
```

## Tidy choccies
```{r}
choc_long <- choc %>%
  melt(id = 1)
```
```{r, eval = F}
head(choc_long)
```
```{r, echo = F}
kable(head(choc_long))
```

## And rename
* We can rename like I did before, with
```{r, eval = F}
names(choc_long)[c(2,3)] <- c('Choccy', 'Prop')
```
* or we can use dplyr::rename
```{r}
choc_long <- choc_long %>%
  rename(Choccy = variable, Prop = value)
```

## Renaming
* Which gives us:
```{r, eval = F}
head(choc_long)
```
```{r, echo = F}
kable(head(choc_long))
```

## Plotting our results
* Just like othe*R* things, there is more than one way to do this
* The two main ways are base R (like our plot() function earlier) and ggplot2
* ggplot2 is basically awesome
* There are other underrated alternatives like lattice and ggplot2::qplot

## Base R
* really easy to use, horrible to get perfect
* Great for quick summaries
```{r}
boxplot(Prop ~ Choccy, data = choc_long)
```

## ggplot2
* ggplot is fully customisable and well-documented (and pretty)
* But we need to learn the syntax
* It's done by layers, where we add a layer each time
* Start by defining ggplot

## ggplot2
```{r}
library(ggplot2)
cplot <- ggplot(data = choc_long)
```

## ggplot2
* We need to add aesthetics to make sure it maps properly
```{r}
cplot <- ggplot(data = choc_long,
                aes(x = Choccy, y = Prop))
```

## ggplot2
* Then we add our layers
* We can do this with either stat_summary() or geom_xxx()
* In most cases stat_summary is what you want
* Except boxplots or histograms basically!

## ggplot2
```{r}
cplot + geom_boxplot()
```

## How about by group?
```{r}
cplot <- ggplot(data = choc_long,
                aes(x = Group, y = Prop)) +
  geom_boxplot()
cplot
```

## What if we want more???
* We can overwrite the aesthetics in additional layers if we want to
* This is super useful for plotting things like model fit vs data
* Gives us a lot of flexibility

## Adding layers
```{r}
cplot + stat_summary(aes(colour = Choccy, shape = Choccy),
                     geom = 'point', fun.y = 'mean')
```

## Adding layers
```{r}
cplot + stat_summary(aes(colour = Choccy, shape = Choccy),
                     geom = 'point', fun.y = 'mean',
                     size = 4)
```

## Adding layers
```{r}
cplot + stat_summary(aes(colour = Choccy, shape = Choccy),
                     geom = 'point', fun.y = 'mean',
                     size = 4) + 
  theme_dark()
```

## Adding layers
```{r}
cplot + stat_summary(aes(colour = Choccy, shape = Choccy),
                     geom = 'point', fun.y = 'mean',
                     size = 4) + 
  theme_classic()
```

## Modern plotting styles
```{r}
cplot <- ggplot(data = choc_long,
                aes(x = Group, y = Prop)) +
  geom_violin() + theme_classic()
cplot
```

## Modern plotting styles
```{r}
cplot <- ggplot(data = choc_long,
                aes(x = Group, y = Prop, colour = Group, fill = Group)) +
  geom_violin() + theme_classic()
cplot
```

## Great data visualisation
```{r}
cplot <- ggplot(data = choc_long,
                aes(x = Group, y = Prop, colour = Group, fill = Group)) +
  geom_violin(alpha = 0.2) + 
  stat_summary(geom = 'pointrange', fun.data = 'mean_se') + 
  theme_classic() +
  ylab('Proportion') +
  xlab('Research Group')
```

## Great data visualisation
```{r}
cplot
```


## Student's Sleep Data
* In 1905, Cushny and Peebles looked at the effect of two soporific drugs on 10 patients
* This was famously re-analyzed by *Student* in 1908, introducing what we now know as the *t*-test
* Fun fact, *Student* was William Sealy Gosset, who was working for Guinness at the time, and the test was used to measure the quality of the beer
* This is a built-in dataset in R

## Student's Sleep Data
```{r}
sleep <- sleep
```
```{r, eval = F}
head(sleep)
```
```{r, echo = F}
kable(head(sleep))
```

## Student's sleep data
* *extra* is the increase in hours of sleep
* *group* is the drug given (poorly named)
* *ID* is the patient ID
* How would we describe this???

## Check our data
* As mentioned last week, you should always check the type of data
* You can do this by hovering over each column in View(sleep)
* What are we looking at?

## T-test in R
* Each function has a default behavior
* t.test(x, y = NULL,
       alternative = c("two.sided", "less", "greater"),
       mu = 0, paired = FALSE, var.equal = FALSE,
       conf.level = 0.95, ...)
* This example appears in the official R examples, and they forget to pair the data!
       
## T-test in R
* We want to use the function interface almost all of the time, unless it is comparing to mu
* That is: y ~ x - remember we used that in dcast()
```{r}
t.test(extra ~ group, paired = T, data = sleep)
```

## Plot our results
* What kind of data visualisation is appropriate here?

## Plot our results
* What kind of data visualisation is appropriate here?
* I'm going to say a boxplot, partly because I don't like barplots
* You can always play with other options

## Plot our results
```{r}
splot <- ggplot(data = sleep,
                aes(x = group, y = extra)) +
  geom_boxplot() +
  theme_classic()
```

## Plot our results
```{r}
splot
```

## Add significance marks?
```{r}
splot + geom_text(x = 1.5, y = 4.5, label = '**', size = 10)
```

## Plant growth data
* Results of an experiment comparing weight of plants under 3 conditions
```{r}
PlantGrowth <- PlantGrowth
```
```{r, eval=F}
head(PlantGrowth)
```
```{r, echo = F}
kable(head(PlantGrowth))
```

## Plant growth data
* Three factors, so we need to do an anova
* Important to note that an anova *is* a linear regression
* Again we use formula syntax
* We'll come back to that...

## Doing a linear model
* There are pure anova ways of doing this in other packages
* For us, it is easiest to run it just like a linear model
* We can use the same syntax as in a *t*-test
```{r}
pgmod <- lm(weight ~ group, data = PlantGrowth)
```

## Linear model
```{r}
summary(pgmod)
```

## Anova table
* Rather than contrasting, we may wish to see the anova table
```{r}
anova(pgmod)
```

## Plotting
* Can you plot these data?

## Plotting
```{r}
pgplot <- ggplot(PlantGrowth, 
                 aes(x = group, y = weight, fill = group)) +
  geom_boxplot(alpha = 0.4) +
  theme_classic()
```

## Plotting
```{r}
pgplot
```